Automatic Matching of Bullet Lands

### 2015-12-30 imaging-paper.Rnw
dotplot instead of table

Besides a probabilistic quantification of matches, we get from a random forest an assessment of the importance of each of the features derived from the bullets' 3d topological surface measurements. Figure~\ref{fig:importance} shows an overview of the importance of each variable measured as the mean decrease in the gini index when the variable in question is included in a tree.
\begin{figure}
<<echo=FALSE, fig.width=7, fig.height=1.5, out.width='.7\\textwidth'>>=
imp <- data.frame(importance(rtrees))
imp$Variable <- row.names(imp)
imp$Variable[c(7,8)] <- c("#matches", "#non-matches")
imp1 <- subset(imp, Variable != c("x1", "sd.D"))
imp2 <- subset(imp, Variable %in% c("x1", "sd.D"))
qplot(MeanDecreaseGini,1, data=imp, geom="point") + 
  geom_text(aes(y=1.1, label=Variable), angle=45, hjust=0, data=imp1) + 
  geom_text(aes(y=0.9, label=Variable), angle=-45, hjust=0, data=imp2) + 
  ylim(c(0.5, 2.5)) + theme_bw() + ylab("") +
  theme(axis.ticks.y=element_blank(), axis.text.y=element_blank(),
        panel.grid.major.y=element_blank(), panel.grid.minor.y=element_blank()) +
  xlab("Importance (mean decrease in Gini index)")
@
\caption{\label{fig:importance}Importance of features in the random forest. Importance is measured in terms of mean decrease in gini index when including the variable in a decision tree.}
\end{figure}
\begin{table}[tbhp]
\caption{\label{tab:importance}Table of  features derived from bullet image ordered by importance in predicting matches. Importance is measured in terms of mean decrease in gini index when including the variable in a decision tree. Averages (and standard deviations) for known matches (KM) and known non-matches (KNM) are shown in the last four columns.}
\centering
<<importance, echo=FALSE, results='asis', warning=FALSE>>=
imp <- data.frame(importance(rtrees))
imp$Variable <- row.names(imp)
imp <- imp[order(-imp$MeanDecreaseGini),]
names(imp)[1] <- "Importance"
imp <- imp[,2:1]
imp$meanMatch <- sapply(imp$Variable, function(var) {
  mean(subset(bstats, match)[,var])
})
imp$sdMatch <- sapply(imp$Variable, function(var) {
  sd(subset(bstats, match)[,var])
})
imp$meanNonMatch <- sapply(imp$Variable, function(var) {
  mean(subset(bstats, !match)[,var])
})
imp$sdNonMatch <- sapply(imp$Variable, function(var) {
  sd(subset(bstats, !match)[,var])
})
imp$meanMatch <- round(imp$meanMatch,1)
imp$sdMatch <- sprintf("(%5.2f)", imp$sdMatch)
imp$meanNonMatch <- round(imp$meanNonMatch,1)
imp$sdNonMatch <- sprintf("(%7.2f)", imp$sdNonMatch)
row.names(imp) <- 1:nrow(imp)
imp$Variable[c(2,4)] <- c("#matches", "#non-matches")
names(imp)[-(1:2)] <- c("KM", "(sd)", "KNM", "(sd)")
print(xtable(imp, align="clrrd{3.2}rd{3.2}", digits=c(0,0,1,1,2,1,2)), 
      floating = FALSE)
@
\end{table}

>>>>

Besides a probabilistic quantification of matches, we get from a random forest an assessment of the importance of each of the features derived from the bullets' 3d topological surface measurements. Figure~\ref{fig:importance} shows an overview of the importance of each variable measured as the mean decrease in the gini index when the variable in question is included in a tree.
\begin{figure}
<<echo=FALSE, fig.width=7, fig.height=1.5, out.width='.7\\textwidth'>>=
imp <- data.frame(importance(rtrees))
imp$Variable <- row.names(imp)
imp$Variable[c(7,8)] <- c("#matches", "#non-matches")
imp1 <- subset(imp, !(Variable %in% c("x1", "sd.D")))
imp2 <- subset(imp, Variable %in% c("x1", "sd.D"))
qplot(MeanDecreaseGini,1, data=imp, geom="point") + 
  geom_text(aes(y=1.1, label=Variable), angle=45, hjust=0, data=imp1) + 
  geom_text(aes(y=0.9, label=Variable), angle=-45, hjust=0, data=imp2) + 
  ylim(c(0, 2.75)) + theme_bw() + ylab("") +
  theme(axis.ticks.y=element_blank(), axis.text.y=element_blank(),
        panel.grid.major.y=element_blank(), panel.grid.minor.y=element_blank()) +
  xlab("Importance (mean decrease in Gini index)")
@
\caption{\label{fig:importance}Importance of features in the random forest. Importance is measured in terms of mean decrease in gini index when including the variable in a decision tree.}
\end{figure}
%' \begin{table}[tbhp]
%' \caption{\label{tab:importance}Table of  features derived from bullet image ordered by importance in predicting matches. Importance is measured in terms of mean decrease in gini index when including the variable in a decision tree. Averages (and standard deviations) for known matches (KM) and known non-matches (KNM) are shown in the last four columns.}
%' \centering
%' <<importance, echo=FALSE, results='asis', warning=FALSE>>=
%' imp <- data.frame(importance(rtrees))
%' imp$Variable <- row.names(imp)
%' imp <- imp[order(-imp$MeanDecreaseGini),]
%' names(imp)[1] <- "Importance"
%' imp <- imp[,2:1]
%' imp$meanMatch <- sapply(imp$Variable, function(var) {
%'   mean(subset(bstats, match)[,var])
%' })
%' imp$sdMatch <- sapply(imp$Variable, function(var) {
%'   sd(subset(bstats, match)[,var])
%' })
%' imp$meanNonMatch <- sapply(imp$Variable, function(var) {
%'   mean(subset(bstats, !match)[,var])
%' })
%' imp$sdNonMatch <- sapply(imp$Variable, function(var) {
%'   sd(subset(bstats, !match)[,var])
%' })
%' imp$meanMatch <- round(imp$meanMatch,1)
%' imp$sdMatch <- sprintf("(%5.2f)", imp$sdMatch)
%' imp$meanNonMatch <- round(imp$meanNonMatch,1)
%' imp$sdNonMatch <- sprintf("(%7.2f)", imp$sdNonMatch)
%' row.names(imp) <- 1:nrow(imp)
%' imp$Variable[c(2,4)] <- c("#matches", "#non-matches")
%' names(imp)[-(1:2)] <- c("KM", "(sd)", "KNM", "(sd)")
%' print(xtable(imp, align="clrrd{3.2}rd{3.2}", digits=c(0,0,1,1,2,1,2)), 
%'       floating = FALSE)
%' @
%' \end{table}
